\LARGE \bf Multi-Arm Bandit Models for 2D Sample Based Grasp Planning with Uncertainty [v15 March 15, 2015 ]

### 2015-03-16 paper.tex
TIdied up a few cloud statements and re-added commits from v15

\begin{abstract}
For applications such as warehouse order fulfillment, robot grasps must be robust to uncertainty arising from sensing, shape, mechanics, and control.  One way to achieve this is to evaluate the performance of candidate grasps by sampling perturbations in shape, pose, and control, computing the probability of force closure for each candidate to identify the grasp with the highest expected quality.  This can be computationally demanding and prior work has turned to cloud based computing to solve it. To improve efficiency, we consider how Multi-Armed Bandit (MAB) models for optimizing sequential decisions can be applied in this context. We formulate robust grasp planning as a MAB problem and evaluate grasp planning convergence time using 100 object shapes randomly selected from the Brown Vision 2D Lab Dataset with 1000 uniformly distributed candidate grasp angles. We consider the case where shape uncertainty is represented as a Gaussian process implicit surface (GPIS) and there is Gaussian uncertainty in pose, gripper approach angle, and coefficient of friction.  Non-MAB methods, uniform allocation and iterative pruning, converge slowly.  In contrast, the MAB methods we consider, Thompson Sampling and the Gittins Index, were more efficient in experiments, converging to within 3\% of the optimal grasp approximately 5x faster than non-MAB methods.
\end{abstract}

@@

\begin{figure}%
    \centering
     \subfloat{{\includegraphics[scale=0.25]{matlab_figures/pfc.eps} }}%

@@

One common measure of quality is force closure, the ability to resist external forces and torques in arbitrary directions~\cite{li1988task}. To cope with uncertainty, recent work has explored computing the probability of  force closure given uncertainty in pose~\cite{christopoulos2007handling, kim2012physically, weisz2012pose} and object shape~\cite{kehoe2012estimating, mahler2015gp}.
One way to compute the probability of force closure is using Monte-Carlo integration over sample perturbations in the uncertain quantities\cite{christopoulos2007handling, kim2012physically, weisz2012pose, kehoe2012toward}. However, performing Monte-Carlo integration for each candidate grasp in a set is computationally expensive.  Past work has looked at using parallel programming in the cloud to overcome this hurdle \cite{kehoe2012estimating,kehoe2012toward,kehoe2015survey} . We show that the Multi-armed Bandit (MAB) model improves upon this brute force approach and could hopefully lead to more efficient cloud based algorithms.

@@

We compare the performance of Thompson sampling and Gittins indices, two popular algorithms for solving the MAB problem, against uniform allocation and an adaptive sampling method known as iterative pruning, which iteratively reduces the set of candidate grasps  based on sample mean~\cite{kehoe2012toward}. In the task of finding grasps with high probability of force closure.
We compare on the Brown Vision 2D Dataset, a dataset of 2D planar objects~\cite{brown, christopoulos2007handling}.
Our initial results in simulation show that Thompson Sampling, a MAB algorithm, required approximately $5\times$ fewer samples than iterative pruning to plan a grasp within $3\%$ of the estimated highest probability of force closure in the set of 1000 randomly selected grasps averaged over 100 objects.

@@

A traditional MAB example is a gambler with $K$ independent one-armed bandits, also known as slot machines.
When an arm is played (or ``pulled'' in the literature), it returns an amount of money from a fixed reward distribution that is unknown to the gambler.
The goal of the gambler is to come up with a method to maximize the sum of average rewards over all pulls.
If the gambler knew the machine with the highest expected reward, the gambler would only pull that arm.
However, since the reward distributions are unknown, a successful gambler needs to trade off exploiting the arms that currently yields the highest expected reward and exploring new arms.
Developing a policy that successfully trades between exploration and exploitation reward has been the focus of extensive research since the problem formulation \cite{bubeck2009pure}, \cite{robbins1985some}, \cite{bergemann2006bandit}.

@@

The solution to the optimization problem assigns each state an index $v(x_{k,t})$.
The indices can then be used to form a policy, where at each timestep the agent selects the arm, $k_t$ where $k_t= \underset{1 \leq k \leq K}{\mbox{argmax }} v(x_{k,t})$.  Traditionally, the indices are computed offline using a variety of methods \cite{weber1992gittins}, we chose to use the restart method proposed by Katehakis et al. \cite{katehakis1987multi} due to its ability to be implemented in a dynamic programming fashion.

@@

\subsection{Candidate Grasp Model}
The grasp model is illustrated in Fig.~\ref{fig:grasp_model}. We formulate the MAB problem for planar objects using parallel-jaw grippers. Similar to~\cite{mahler2015gp}, we parameterize a grasp using a {\it grasp axis}, the axis of approach for two jaws,  with jaws of width $w_j \in \mathbb{R}$ and a maximum width $w_g \in \mathbb{R}$. The two location of the jaws can be specified as $\mathbf{j}_1,\mathbf{j}_2 \in \mathbb{R}^2$, where $||\mathbf{j}_1 - \mathbf{j}_2||_2 \leq w_g$. We define a grasp consisting of the tuple $\Gamma = \lbrace \mathbf{j}_1, \mathbf{j}_2 \rbrace$.

@@

\subsection{Grasp Quality}\label{sec:grasp_sample}
We measure the quality of grasp using the probability of force closure~\cite{weisz2012pose, kim2012physically, kehoe2012estimating, kehoe2012toward} given a grasp $\Gamma$. Force closure is a binary-valued quantity $F$ that is 1 if the grasp can resist wrenches in arbitrary directions and 0 otherwise.
Let $\mW \in \mathbb{R}^3$ denote the contact wrenches derived from contact locations $\bc_1,\bc_2$, normals $\bn_1,\bn_2$, friction coefficient $\mu$, and center of mass $\bz$ for a given grasp and shape.
If the origin lies within the convex hull of $\mW$, then the grasp is in force closure~\cite{li1988task}.
We rank grasps using the probability of force closure given uncertainty in shape, pose, robot approach, and friction coefficient~\cite{christopoulos2007handling, kehoe2012toward}:

@@

To summarize the algorithm terminates when the following condition is met

\vspace{-2ex}
\begin{align}\label{eq:stopping_condition}
\lbrace t \geq T_s \: \mbox{OR} \: |B(0.025,\alpha_{\bar{k},t},\beta_{\bar{k},t}) -  B(0.975,\alpha_{\bar{k},t},\beta_{\bar{k},t})| \leq \epsilon \rbrace
\end{align}
\noindent where the $\bar{k}$ grasp refers to the grasp with the highest estimated $P_F$ at time $t$.

\section{Simulation Experiments}
We used the Brown Vision Lab 2D dataset \cite{brown}, a database of 2D planar objects, as in \cite{christopoulos2007handling}.
GPIS models of 3 example objects are visualized using the GPIS-Blur method of~\cite{mahler2015gp} in Figure \ref{fig:shape_samples}.
We downsampled the silhouette by a factor of 2 to create a 40 x 40 occupancy map, which holds 1 if the point cloud was observed and 0 if it was not observed, and a measurement noise map, which holds the variance 0-mean noise added to the SDF values.
The measurement noise were assigned uniformly at random to the grid cells.
We then construct a GPIS using the same method proposed in~\cite{mahler2015gp}.
For illustrative purposes, the noise in approach, pose, and friction coefficient were set to the following variances $\sigma_{\psi} = 0.2$ rads, $\sigma_{\by} = 3$ grid cells, $\sigma_{\mu} = 0.4$, $\sigma_{\phi} = 0.3$ rads and $\sigma_{\bt} = 3$ grid cells.
We also performed experiments for the case of two hard contacts in 2-D.
We drew random grasps $\Gamma$ by sampling the angle of grasp axis around a circle with radius $\sqrt{2} M$, where $M$ is the dimension of the workspace, and then sampling the circle's origin. All experiments were run on machine with OS X with a 2.7 GHz Intel core i7 processor and 16 GB 1600 MHz memory in Matlab 2014a.

\subsection{Multi-Armed Bandit Experiments}
For our experiments we look at selecting the best grasp out of a size of $|G| = 1000$. We draw samples from our graphical model using the technique described in Sec.  \ref{sec:grasp_sample}. In Fig. \ref{fig:grasp_quality}, we plotted the probability of force closure, $P_F$, of the grasp chosen vs. stopping time $T_s$ averaged over 100 randomly selected shapes in the Brown Vision Lab 2D dataset and for the grasps planned by Thompson sampling, Gittins indices, iterative pruning \cite{kehoe2012toward} and uniform allocation). Uniform allocation selects a grasp at random from the set to sample next, thus does not use any prior information. iterative pruning prunes grasps every 1000 iterations based on lowest sample mean and removes 10$\%$ of the current grasp set.  We set the discount factor $\gamma =0.98$ for Gittins because that was the highest we could compute the indices for in a feasible amount of time.

@@

\begin{figure*}%
\label{fig:shape_examples}
    \centering
    %\includegraphics[width=16.5cm]{figures/shapes.jpg}%
    \caption{Grasps with the highest estimated probabilty of force closure $P_F$ after 10,000 samples using uniform allocation, iterative pruning, Gittins indices, and Thompson sampling on three objects from the Brown Vision 2D Dataset. For reference, we also show the grasp with highest $P_F$ after brute force evaluation using 100,000 samples. The candidate grasp set was of size $|\mG|=1000$. We visualize the GPIS representation of the object shape uncertainty using GPIS-Blur~\cite{mahler2015gp}, which causes more uncertain areas to appear more blurry.  On objects A and B Thompson sampling finds the best grasp after 10,000 samples, but for object C the grasp selected has $2\%$ lower $P_F$.}%
    \label{fig:shape_samples}%
\end{figure*}

@@

As illustrated in Fig. \ref{fig:worst_case}, all the methods are affected by worst case performance. It would appear that when the observations are misleading the best thing to do is simply uniform allocation of grasp samples. Interestingly, Thompson sampling appears to to continue to improvement while Gittins indices and iterative pruning are stuck in local optimal.

\begin{figure}[t!]
\centering
\includegraphics[scale=0.25]{matlab_figures/worst_case.eps}
\caption{ \footnotesize 
	Number of samples versus the probability of force closure of the best estimated grasp after $k$ samples $P_F(\Gamma_{\bar{k}})$ using uniform allocation, iterative pruning, Gittins indices, and Thompson Sampling over 1,000 candidate grasps when samples are ordered such that many perturbations that do result in force closure are given to the algorithm initially.
	The normalized $P_F$ is the ratio of the best estimated grasp at iteration $k$, $P_F(\Gamma_{\bar{k}})$, to the highest $P_F$ in the candidate grasp set $P_F(\Gamma^*)$ averaged over 100 independent runs on randomly selected objects from the Brown Vision 2D Dataset \cite{brown}.
 	The highest quality grasp was determined by brute force search over all candidate grasps (which required 10x  more time than any of these methods \cite{kehoe2012toward}).
 	The results suggest that when samples are misleading the best policy is uniform allocation. 	However, Thompson sampling appears to continue making progress whereas iterative pruning and Gittins indices do not.}
\vspace*{-10pt}
\label{fig:worst_case}
\end{figure}

@@

In future work, we plan to scale our method to 3D objects. However, this could substantially increase the number of candidate grasps. One option to solving this problem is to use the cloud to solve subsets of the MAB problem in parralel. By dividing up the MAB into batches of arms and solving each batch in parallel, we can hopefully maintain similar performance as in this paper.

>>>>

\begin{abstract}
For applications such as warehouse order fulfillment, robot grasps must be robust to uncertainty arising from sensing, shape, mechanics, and control.  One way to achieve this is to evaluate the performance of candidate grasps by sampling perturbations in shape, pose, and control, computing the probability of force closure for each candidate to identify the grasp with the highest expected quality. Prior work has turned to cloud based computing because evaluating the quality of each grasp is computationally demanding. To improve efficiency, we consider how Multi-Armed Bandit (MAB) models for optimizing decisions can be applied in this context. We formulate robust grasp planning as a MAB problem and evaluate grasp planning convergence time using 100 object shapes randomly selected from the Brown Vision 2D Lab Dataset with 1000 uniformly distributed candidate grasp angles. We consider the case where shape uncertainty is represented as a Gaussian process implicit surface (GPIS) and there is Gaussian uncertainty in pose, gripper approach angle, and coefficient of friction. 
Uniform allocation and iterative pruning, the non-MAB methods, converge slowly.
In contrast, Thompson Sampling and the Gittins index method, the MAB methods we consider, converged to within 3\% of the optimal grasp approximately 5x faster.
\end{abstract}

@@

\begin{figure}%
    \centering
     \subfloat{{\includegraphics[scale=0.25]{matlab_figures/pfc-01.jpg} }}%

@@

One common measure of quality is force closure, the ability to resist external forces and torques in arbitrary directions~\cite{li1988task}. To cope with uncertainty, recent work has explored computing the probability of  force closure given uncertainty in pose~\cite{christopoulos2007handling, kim2012physically, weisz2012pose} and object shape~\cite{kehoe2012estimating, mahler2015gp}.
One way to compute the probability of force closure is using Monte-Carlo integration over sample perturbations in the uncertain quantities\cite{christopoulos2007handling, kim2012physically, weisz2012pose, kehoe2012toward}. However, performing Monte-Carlo integration for each candidate grasp in a set is computationally expensive.  Past work has looked at using parallel programming in the cloud to overcome this hurdle \cite{kehoe2012estimating,kehoe2012toward,kehoe2015survey}.
In this work, we aim to extend these cloud-based methods by reducing the number of samples needed to converge to a high-quality grasp.

@@

We compare the performance of Thompson sampling and Gittins indices, two popular algorithms for solving the MAB problem, against uniform allocation and an adaptive sampling method known as iterative pruning, which iteratively reduces the set of candidate grasps  based on sample mean~\cite{kehoe2012toward}. In the task of finding grasps with high probability of force closure.
We compare on the Brown Vision 2D Dataset, a dataset of 2D planar objects~\cite{brown, christopoulos2007handling}.
We demonstrate performance on a local implementation. However, our methods are parallelizable by dividing the arms into $M$ subset,where $M$ is the number of available cores, and solving each MAB problem separately\cite{glazebrook2000index}. 
Our initial results in simulation show that Thompson Sampling, a MAB algorithm, required approximately $5\times$ fewer samples than iterative pruning to plan a grasp within $3\%$ of the estimated highest probability of force closure in the set of 1000 randomly selected grasps averaged over 100 objects.

@@

A traditional MAB example is a gambler with $K$ independent one-armed bandits, also known as slot machines.
When an arm is played (or ``pulled'' in the literature), it returns an amount of money from a fixed reward distribution that is unknown to the gambler.
The goal of the gambler is to come up with a method to maximize the sum of average rewards over all pulls.
If the gambler knew the machine with the highest expected reward, the gambler would only pull that arm.
However, since the reward distributions are unknown, a successful gambler needs to trade off exploiting the arms that currently yields the highest expected reward and exploring new arms.
Developing a policy that successfully trades between exploration and exploitation reward has been the focus of extensive research since the problem formulation  ~\cite{bergemann2006bandit},\cite{bubeck2009pure}, \cite{robbins1985some}.

@@

The solution to the optimization problem assigns each state an index $v(x_{k,t})$.
The indices can then be used to form a policy, where at each timestep the agent selects the arm $k_t$ where $k_t= \underset{1 \leq k \leq K}{\mbox{argmax }} v(x_{k,t})$.  Consult \cite{gittins2011multi} for a more detailed analysis of the Gittins method. The indices can be  computed offline using a variety of methods \cite{weber1992gittins}; we chose to use the restart method proposed by Katehakis et al. \cite{katehakis1987multi} because it can be implemented in a dynamic programming fashion.

@@

\subsection{Candidate Grasp Model}
The grasp model is illustrated in Fig.~\ref{fig:grasp_model}. We formulate the MAB problem for planar objects using parallel-jaw grippers as modeled in Fig.~\ref{fig:grasp_model}. Similar to~\cite{mahler2015gp}, we parameterize a grasp using a {\it grasp axis}, the axis of approach for two jaws,  with jaws of width $w_j \in \mathbb{R}$ and a maximum width $w_g \in \mathbb{R}$. The two location of the jaws can be specified as $\mathbf{j}_1,\mathbf{j}_2 \in \mathbb{R}^2$, where $||\mathbf{j}_1 - \mathbf{j}_2||_2 \leq w_g$. We define a grasp consisting of the tuple $\Gamma = \lbrace \mathbf{j}_1, \mathbf{j}_2 \rbrace$.

@@

\subsection{Grasp Quality}\label{sec:grasp_sample}
We measure the quality of grasp using the probability of force closure~\cite{kehoe2012estimating, kehoe2012toward, kim2012physically,weisz2012pose} given a grasp $\Gamma$. Force closure is a binary-valued quantity $F$ that is 1 if the grasp can resist wrenches in arbitrary directions and 0 otherwise.
Let $\mW \in \mathbb{R}^3$ denote the contact wrenches derived from contact locations $\bc_1,\bc_2$, normals $\bn_1,\bn_2$, friction coefficient $\mu$, and center of mass $\bz$ for a given grasp and shape.
If the origin lies within the convex hull of $\mW$, then the grasp is in force closure~\cite{li1988task}.
We rank grasps using the probability of force closure given uncertainty in shape, pose, robot approach, and friction coefficient~\cite{christopoulos2007handling, kehoe2012toward}:

@@

To summarize the algorithm terminates and returns $\bar{k}$, or the grasp that has the highest estimated $P_F$ when the following condition is met:

\vspace{-2ex}
\begin{align}\label{eq:stopping_condition}
\lbrace t \geq T_s \: \mbox{OR} \: |B(0.025,\alpha_{\bar{k},t},\beta_{\bar{k},t}) -  B(0.975,\alpha_{\bar{k},t},\beta_{\bar{k},t})| \leq \epsilon \rbrace.
\end{align}

\section{Simulation Experiments}
We used the Brown Vision Lab 2D dataset \cite{brown}, a database of 2D planar objects, as in \cite{christopoulos2007handling}.
GPIS models of 3 example objects are visualized using the GPIS-Blur method of~\cite{mahler2015gp} in Figure \ref{fig:shape_samples}.
We downsampled the silhouette by a factor of 2 to create a 40 x 40 occupancy map, which holds 1 if the point cloud was observed and 0 if it was not observed, and a measurement noise map, which holds the variance 0-mean noise added to the SDF values.
The measurement noise were assigned uniformly at random to the SDF.
We then construct a GPIS using the same method proposed in~\cite{mahler2015gp}.
For illustrative purposes, the noise in approach, pose, and friction coefficient were set to the following variances $\sigma_{\psi} = 0.2$ rads, $\sigma_{\by} = 3$ grid cells, $\sigma_{\mu} = 0.4$, $\sigma_{\phi} = 0.3$ rads and $\sigma_{\bt} = 3$ grid cells.
We also performed experiments for the case of two hard contacts in 2-D.
We drew random grasps $\Gamma$ by sampling the angle of grasp axis around a circle with radius $\sqrt{2} M$, where $M$ is the dimension of the workspace, and then sampling the circle's origin. All experiments were run on machine with OS X with a 2.7 GHz Intel core i7 processor and 16 GB 1600 MHz memory in Matlab 2014a.

\subsection{Multi-Armed Bandit Experiments}
For our experiments we look at selecting the best grasp out of a size of $|G| = 1000$. We draw samples from our graphical model using the technique described in Sec.  \ref{sec:grasp_sample}. We averaged over 100 randomly selected shapes in the Brown Vision Lab 2D dataset and for the grasps planned by Thompson sampling, Gittins indices, iterative pruning \cite{kehoe2012toward} and uniform allocation). Uniform allocation selects a grasp at random from the set to sample next, thus does not use any prior information. iterative pruning prunes grasps every 1000 iterations based on lowest sample mean and removes 10$\%$ of the current grasp set.  We set the discount factor $\gamma =0.98$ for Gittins because that was the highest we could compute the indices for in a feasible amount of time.

@@

\begin{figure*}%
\label{fig:shape_examples}
    \centering
    \includegraphics[width=16.5cm]{figures/shapes.jpg}%
    \caption{Grasps with the highest estimated probabilty of force closure $P_F$ after 10,000 samples using uniform allocation, iterative pruning, Gittins indices, and Thompson sampling on three objects from the Brown Vision 2D Dataset. For reference, we also show the grasp with highest $P_F$ after brute force evaluation using 100,000 samples. The candidate grasp set was of size $|\mG|=1000$. We visualize the GPIS representation of the object shape uncertainty using GPIS-Blur~\cite{mahler2015gp}, which causes more uncertain areas to appear more blurry.  On objects A and B Thompson sampling finds the best grasp after 10,000 samples, but for object C the grasp selected has $2\%$ lower $P_F$.}%
    \label{fig:shape_samples}%
\end{figure*}

@@

As illustrated in Fig. \ref{fig:worst_case}, all the methods are affected by worst case performance. It would appear that when the observations are misleading the best thing to do is simply uniform allocation of grasp samples. Interestingly, Thompson sampling appears to to continue to improvement while Gittins indices and iterative pruning do not continue to make progress.

\begin{figure}[t!]
\centering
\includegraphics[scale=0.25]{matlab_figures/worst_case-02.jpg}
\caption{ \footnotesize 
	Number of samples versus the probability of force closure of the best estimated grasp after $k$ samples $P_F(\Gamma_{\bar{k}})$ using uniform allocation, iterative pruning, Gittins indices, and Thompson Sampling over 1,000 candidate grasps when samples are ordered such that many perturbations that do result in force closure are given to the algorithm initially.
	The normalized $P_F$ is the ratio of the best estimated grasp at iteration $k$, $P_F(\Gamma_{\bar{k}})$, to the highest $P_F$ in the candidate grasp set $P_F(\Gamma^*)$ averaged over 100 independent runs on randomly selected objects from the Brown Vision 2D Dataset \cite{brown}.
 	The highest quality grasp was determined by brute force search over all candidate grasps (which required 10x  more time than any of these methods \cite{kehoe2012toward}).
 	The results suggest that when samples are misleading the best policy is uniform allocation. 	However, Thompson sampling appears to continue making progress whereas iterative pruning and Gittins indices do not.}
\vspace*{-10pt}
\label{fig:worst_case}
\end{figure}

@@

In future work, we plan to scale our method to 3D objects. This could substantially increase the number of candidate grasps, further motivating the use of cloud computing. Glazebrook and Wilkinson showed that the Gittins index method could be parallelized by simply dividing the arms into $M$ subsets, where $M$ is the number of cores, and solving each MAB separately. \cite{glazebrook2000index}. A similar method could also be done for Thompson sampling. Another promising scheme for parrallelizing the MAB, is to have the algorithm not sample one arm at each algorithm but $M$. We will explore both of these approaches in future work.
